Random Forests", OAs

### 2023-12-21 src/oa_RF.Rmd
refined model LA
local minor changes
refined model OA

```{r}
dist.retail <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()

@@

# Minimum distance 
dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)

@@

@@

@@

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(oa11cd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("oa11cd" = "oa11cd"))

df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                         rename(n.nearest.city = n), 
                       by = c("nearest.city.oa11cd" = "oa11cd",
                              "year" = "year"))

# Lookup: OA to regions
path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)

@@

# df %>% ungroup() %>% distinct(RGN11NM)
# sapply(df, function(x) sum(is.na(x)))

@@

# For growth:
#df <- df %>% mutate(growth = log(n/n.lag))

@@

# in.path <- paste0(path, "/data/temp/df_for_oa_rf_old.csv")
# df.compare <- read_csv(in.path)

# out.path <- paste0(path, "/data/temp/for_oa_rf.RData")
# save.image(file = out.path)
# load(out.path)

df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
                        "Leicester", "nearest.city.oa11cd"))

# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df <- read_csv(out.path)
# df %>% write_csv(file = out.path)
```

@@

```{r}
df <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
  mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
  ungroup() %>% 
  dplyr::select(-help)

df <- df %>% filter(year > 1996)

#df <- df %>% dplyr::select(-n.lag)
```

@@

@@

start.time <- Sys.time()
model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

# stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

@@

#folds
k <- 3 # 7.09h for k=3

@@

detectCores()
cl <- makePSOCKcluster(2)
registerDoParallel(cl)

>>>>

```{r}
dist.retail <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail<- round((dist.retail/1000),2) %>% 
  as_tibble()

@@

# This efficiently returns the name of the column with the shortest distance
dist.retail <- dist.retail %>% mutate(dist.retail.name = names(.)[max.col(.[2:6424]*-1)+1L])

# dist.retail_ <- dist.retail %>% slice_sample(n=100)
# dist.retail_[,6425:6426]
# dist.retail_ <- transform(dist.retail_, dist.retail = do.call(pmin, dist.retail_[2:6424])) 
# sum(is.na(dist.retail_$dist.retail.name))

# Minimum distance 
dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[2:6424]))

dist.retail <- dist.retail  %>% dplyr::select(oa11cd, dist.retail.name, dist.retail)
# 232,296 
#sapply(dist.retail, function(x) sum(is.na(x))) #0

#spatial join retail centroids into OA
retail.major.cetres.help <- st_centroid(retail) %>% 
  st_transform(4326)

# lookup obj with retail id and OA id
help <- st_join(retail.major.cetres.help, oa.gb) %>% 
  as_tibble %>% 
  mutate(dist.retail.name = row_number(),
         dist.retail.name = paste0("V", dist.retail.name)) %>% 
  dplyr::select(id, dist.retail.name) %>% 
  rename(oa11cd.retail = id)

dist.retail <- dist.retail %>% 
  left_join(help, by = c('dist.retail.name' = 'dist.retail.name')) %>% 
  dplyr::select(-dist.retail.name)
sapply(dist.retail, function(x) sum(is.na(x))) 
# 1 retail centre is missing, because the centroid falls in the water

dist.retail %>% filter(is.na(oa11cd.retail))
missing.retail.centre <- retail.major.cetres.help %>% filter(RC_ID=="RC_EW_24")
st_nearest_feature(missing.retail.centre, oa.gb)
oa.gb[91986,] # missing OA: E00094765

dist.retail <- dist.retail %>%
  mutate(oa11cd.retail = ifelse(is.na(oa11cd.retail), "E00094765", oa11cd.retail))

# dist.retail %>% distinct(dist.retail.name) #6405
# help %>% distinct(dist.retail.name) # 6423 
# dist.retail %>% distinct(oa11cd.retail) #6217 as there are multiple centres in an OA

@@

sapply(df, function(x) sum(is.na(x)))
df %>% filter(is.na(oa11cd.retail)) %>% distinct(oa11cd) #493 rows or 29 OA don't have oa11cd.retail

# df <- df %>% dplyr::select(-dist.retail.x, -oa11cd.retail.x) %>% 
#   rename(dist.retail = dist.retail.y,
#          oa11cd.retail = oa11cd.retail.y)
```

@@

help.retail <- dist.retail

@@

# lags
df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(oa11cd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("oa11cd" = "oa11cd"))

# n nearest city
df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                         rename(n.nearest.city = n), 
                       by = c("nearest.city.oa11cd" = "oa11cd",
                              "year" = "year")) 
# n London 
df <- df %>% left_join(df %>% 
                         ungroup() %>% 
                         filter(oa11cd == "E00176659") %>% 
                         dplyr::select(year, n) %>% 
                         rename(n.London = n)) %>% 
  group_by(oa11cd) %>% 
  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
  ungroup()

# n nearest retail
df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                         rename(n.nearest.retail = n), 
                       by = c("oa11cd.retail" = "oa11cd",
                              "year" = "year")) %>% 
  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))

# Lookup: OA to regions
# NEED TO ADD NI. NOT A PROBLEM FOR NOW AS IT IS FIXED IN THE TEST SECTION
path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)

@@

@@

# For growth:
df <- df %>% mutate(growth = log(n/n.lag),
                    abs.growth = n - n.lag)

# regional identifier
df <- df %>% mutate(help = substr(oa11cd, start = 1, stop = 2)) %>% 
  mutate(RGN11CD = ifelse(is.na(RGN11CD), help, RGN11CD)) %>% 
  ungroup() %>% 
  dplyr::select(-help)

@@

# remove not used columns
df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
                        "Leicester", "nearest.city.oa11cd"))

# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df <- read_csv(out.path)
# df %>% write_csv(file = out.path)

@@

```{r}
london.help <- df %>% group_by(RGN11NM, year) %>% 
  summarise(n.London = sum(n)) %>% 
  filter(RGN11NM == "London") %>%
  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
  ungroup() %>% 
  dplyr::select(-RGN11NM)

#n.nearest.retail IS WRONG. FIX UPSTREAM
df <- df %>% 
  rename(oa11cd.retail = n.nearest.retail) %>% 
  dplyr::select(-n.nearest.retail.lag) %>% 
  left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                         rename(n.nearest.retail = n), 
                       by = c("oa11cd.retail" = "oa11cd",
                              "year" = "year")) %>% 
  mutate(n.nearest.retail.lag = dplyr::lag(n.nearest.retail))

df <- df %>% dplyr::select(-n.London, -n.London.lag) %>% 
  left_join(london.help, by = c('year' = 'year'))

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996) %>% 
  dplyr::select(oa11cd, year, RGN11CD, n, n.l.slag, n.nearest.city.lag, n.nearest.retail.lag, n.London.lag, London, dist, dist.retail) #growth, abs.growth, RGN11NM,

df %>% group_by(year) %>% 
  filter(n == 0) %>% 
  count(n)
# in 2012, 120848 0s

# df %>% dplyr::select(area, dist, dist.retail, London, south, n.slag, n.lag, n.l.slag, n.nearest.city, year) %>% 
#   #sapply(function(x) sum(is.na(x)))
#   summarise(across(where(is.numeric), .fns = 
#                      list(#n = is.na,
#                           min = min,
#                           median = median,
#                           mean = mean,
#                           stdev = sd,
#                           # q25 = ~quantile(., 0.25),
#                           # q75 = ~quantile(., 0.75),
#                           max = max))) %>%
#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))
# 
# df %>% dplyr::select(n.London.lag, n.nearest.city.lag, n.nearest.retail.lag, n.l.slag, year, London, dist, dist.retail) %>% 
#   #sapply(function(x) sum(is.na(x)))
#   summarise(across(where(is.numeric), .fns = 
#                      list(#n = is.na,
#                           min = min,
#                           median = median,
#                           mean = mean,
#                           stdev = sd,
#                           # q25 = ~quantile(., 0.25),
#                           # q75 = ~quantile(., 0.75),
#                           max = max))) %>%
#   pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # 9 after adding South. It should be RHS variables + 2

@@

start.time <- Sys.time()
model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                     n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

# stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken #1.3h

@@

#folds
k <- 10 # 7.09h for k=3

@@

detectCores()
cl <- makePSOCKcluster(8)
registerDoParallel(cl)
